### 2014-01-28 rscripts/manuscript.Rnw
more updates

% define Scala
\lstdefinelanguage{scala}{morekeywords={class,object,trait,extends,with,new,if,while,for,def,val,var,this,case,->,=>},
%otherkeywords={->,=>},
%procnamekeys={def,.},
sensitive=true,
keywordstyle=\color{magenta},
commentstyle=\color{green},
identifierstyle=\color{blue},
%procnamestyle=\color{black},
morecomment=[l]{//},
morecomment=[s]{/*}{*/},
morestring=[b]"}
% Default settings for code listings
\lstset{frame=tb,language=scala,aboveskip=3mm,belowskip=3mm,
showstringspaces=false,columns=flexible,basicstyle={\small\ttfamily}}
% handle tilde characters
\lstset{literate=%
{~}{\url{~}}1
}
% define no break code listing
\lstnewenvironment{code}[1][]%
{
   \noindent
   \minipage{\linewidth} 
   \vspace{0.5\baselineskip}
   \lstset{basicstyle=\ttfamily\footnotesize,frame=single,#1}}
{\endminipage}

@@

\section*{Introduction}
For investigators ranging from anthropologists to law enforcement, the desire to find, study, and modify the behavior groups operating at the edge of the observable is a principle concern.  In the meta-theatre of the ironic, various media outlets have responded to public interest by applying this lens to the activities of intelligence organizations world-wide -- themselves desperate to undertake this very task against other clandestine actors.  Of course the revealation of these activities has broadly sparked the symmetric concern: being able to operate clandestinely in an age of ubiquitous monitoring.  Criminal organizations have long appreciated the value of operating in secret, as have groups subject to State-sponsored abuse or industrial espionage, and time will tell whether or not this spark catches fire in general behavior.

@@

One syntactical aside for the code the below sections: the Scala collections (like \texttt{Set}) define methods for adding and removing elements in terms of the natural mathematical operators (\texttt{+}, \texttt{-}) and for similar operations with groups of elements (\texttt{++}, \texttt{--}).  Similarly, the names for message senders (\texttt{sender}), changing context (\texttt{context become ...}), et cetera are all quite clear.  Given the informative, mixed-natural-and-mathematical-like syntax, we prefer actual code snippets to pseudo-code to illustrate simulation steps.

@@

\begin{code}
  def plotter(
    // a plotter has superiors, collaborators, and subordinates
    // these groups are initially empty
    superiors:Set[Plotter] = Set.empty,
    collaborators:Set[Plotter] = Set.empty,
    subordinates:Set[Plotter] = Set.empty) : Receive = {
      // ... to-be-defined behavior
      case _ => // ignore unmatched messages
  }
\end{code}

We need to define how they adopt and evolve this context.  Since we are not studying the organizational dynamics with this simulation, we can be phenomenologically loose for this part, though this approach means we can add monitoring that activity as part of future detection scenarios with minimal coding.  The simulation runner will send a \texttt{Radicalize} message to the initial cabal, then a \texttt{Collaboration} message to the same group:

\begin{code}
  case class Radicalize(id:Long)
  case class Collaboration(id:Long, group:Set[Plotter])

  def receive = {
    case Radicalize(_) =>
      context become plotter()
    case _ => // ignore other messages
  }

@@

\begin{code}
  case class Subordinates(id:Long, recruits:Set[Plotter])
  case class Recruit(id:Long)
  case class RecruitAck(id:Long) // acknowledge recruitment

  def plotter(...) = {
    case Subordinates(_, recruits) =>
      recruits foreach { recruit =>
        recruit ! Recruit(0)
      }
    case RecruitAck(_) =>
      context become plotter(superiors, collaborators, subordinates + sender)
    // ... other cases
  }

  def receive = {
    case Recruit(id) =>
      context become plotter(superiors = Set(sender))
      sender ! RecruitAck(id)
    // ... other cases
  }
\end{code}

@@

The simulation runner sends \texttt{Time(t:Long)} messages to all actors in the simulation to indicate the march of time.  In general, during each time step a leader may initiate a new \texttt{Bad} message.  Other members of the covert group circulate \texttt{Bad} messages subsequent to receiving a \texttt{Bad} message themselves, with different probabilities for sending messages to superiors, collaborators, and subordinates.  Those probabilities do not depend on the origin of the received message (though this could be a useful mechanical tweak to the model).

Though those probabilities could be coded into the actors

>>>>

% define Scala
\lstdefinelanguage{scala}{morekeywords={package,class,object,trait,extends,with,new,if,while,for,def,val,var,this,case,import,type,->,=>},
%otherkeywords={->,=>},
%procnamekeys={def,.},
sensitive=true,
keywordstyle=\color{magenta},
commentstyle=\color{green},
identifierstyle=\color{blue},
%procnamestyle=\color{black},
morecomment=[l]{//},
morecomment=[s]{/*}{*/},
morestring=[b]"}
% Default settings for code listings
\lstset{frame=tb,language=scala,aboveskip=3mm,belowskip=3mm,
showstringspaces=false,columns=flexible,basicstyle={\small\ttfamily}}
% handle tilde characters
\lstset{literate=%
{~}{\url{~}}1
}
% define no break code listing
\lstnewenvironment{code}[1][]%
{
   \noindent
   \minipage{\linewidth} 
   \vspace{0.5\baselineskip}
   \lstset{basicstyle=\ttfamily\footnotesize,frame=single,#1}}
{\endminipage}

@@

\section*{Introduction}
For investigators ranging from anthropologists to law enforcement, the desire to find, study, and modify the behavior groups operating at the edge of the observable is a principle concern.  In the meta-theatre of the ironic, various media outlets have responded to public interest by applying this lens to the activities of intelligence organizations world-wide -- themselves desperate to undertake this very task against other clandestine actors.  The revealation of these activities has sparked the symmetric concern: being able to operate clandestinely in an age of ubiquitous monitoring.  Criminal organizations have long appreciated the value of operating in secret, as have groups subject to State-sponsored abuse or industrial espionage, and time will tell whether or not this spark catches fire in general behavior.

@@

One syntactical aside for the code the below sections: the Scala collections (like \texttt{Set}) define methods for adding and removing elements in terms of the natural mathematical operators (\texttt{+}, \texttt{-}) and for similar operations with groups of elements (\texttt{++}, \texttt{--}).  Similarly, the names for message senders (\texttt{sender}), changing context (\texttt{context become ...}), et cetera are all quite clear.  Given the informative, mixed-natural-and-mathematical-like syntax, we prefer actual code snippets to pseudo-code to illustrate simulation steps.  However, we have elided some declarations that enable the more informative syntax from the initial appearance in code.

@@

\begin{code}
  def plotter(
    // a plotter has superiors, collaborators, and subordinates
    // these groups are initially empty
    superiors:Plotters = empty,
    collaborators:Plotters = empty,
    subordinates:Plotters = empty
    /*, ... other args */) : Receive = {
      // ... to-be-defined behavior
      case _ => // ignore unmatched messages
  }
\end{code}

We need to define how they adopt and evolve this context.  Since we are not studying the organizational dynamics with this simulation, we can be phenomenologically loose for this part, though this approach means we can add monitoring that activity as part of future detection scenarios with minimal coding.  The simulation runner will send a \texttt{Radicalize} message to the initial cabal, then a \texttt{Collaborate} message to the same group:

\begin{code}
  object Radicalize
  case class Collaborate(group:Plotters)

  def receive = {
    case Radicalize =>
      context become plotter
    case _ => // ignore other messages
  }

@@

\begin{code}
  case class Subordinates(recruits:Plotters)
  object Recruit

  def plotter(...) = {
    case Subordinates(_, recruits) =>
      recruits foreach { recruit =>
        recruit ! Recruit
      }
    // ... other cases
  }

  def receive = {
    case Recruit =>
      context become plotter(superiors = Set(sender))
    // ... other cases
  }
\end{code}

@@

The simulation runner sends \texttt{Time(t:Long)} messages to all actors in the simulation to indicate the march of time (waiting for them to complete all their business for that step and reply with \texttt{TimeAck(t)}).  In general, during each time step a leader may initiate a new \texttt{Bad} message.  Other members of the covert group circulate \texttt{Bad} messages subsequent to receiving a \texttt{Bad} message themselves, with different probabilities for sending messages to superiors, collaborators, and subordinates.  Those probabilities do not depend on the origin of the received message (though this could be a useful mechanical tweak to the model).

Though those probabilities could be coded into the actors as part of their initialization, if they are set and adjusted by events then the actors (i) more accurrately reflect real behavior and (ii) the evolution of individuals occurs via a consistent simulation mechanism.  The former provides a lever for expert insight and empirical study.  The latter makes it easier to produce real-alike data -- i.e., simulation data that has the form of real observations.

In our simplified example, we do not consider how ``real'' events would translate into adjusting a covert member's behavior.  Instead, the simulation runner sends a message, giving us a wedge for later adjustment without frontloading too much model complexity.  This represents the final step for how the plotters evolve in this simulation, so the following repeats the previous snippets with some condensing modifications to the code.  The most notable one is making use of a \texttt{case class} to represent all of the parameters that form a plotter's state.

\begin{code}
package object simactor { // defs to make implementation more natural
  type Plotters = Set[ActorRef]
  def empty : Plotters = Set.empty
  def group(list:ActorRef*) : Plotters = list.toSet
  type Probability = Double
}

object SimActor { // defines common elements for SimActors
  // MESSAGES ABOUT INTERNAL STATE
  object Radicalize                           // become a plotter
  object Recruit                              // work for a plotter
  case class Collaborate(group:Plotters)      // work with other plotters
  case class Subordinates(recruits:Plotters)  // recruit subordinates

  object PType extends Enumeration { // plotter behavior probabilites
    type PType = Value
    val Initiate, // prob. of starting plot comms
        Report,   // prob. of reporting to superiors
    	  Collab,	  // prob. of working w/ peers
    	  Delegate  // prob. of delegating work
    		  = Value
  }
  case class UpdateProb(changes:Map[PType.Value,Probability])

  // MESSAGES ABOUT EXTERNAL ACTIVITY
  object Bad
  object Good

  case class PlotterState( // the plotter state description
    superiors:Plotters = empty,
    collaborators:Plotters = empty,
    subordinates:Plotters = empty,
    probs:Map[PType.Value,Probability] = PType.values.map { (_, 0.0) } toMap
    // default each plotting probability 0.0
  )
}

class SimActor extends Actor {
  import SimActor._

  def receive = {
    case Radicalize =>
      context become plotter()
    case Recruit =>
      context become plotter(PlotterState(superiors = group(sender)))
    case _ => // ignore other messages
  }

  def plotter(ps:PlotterState = PlotterState()) : Receive = {
    import ps._
    def update(newPs:PlotterState) = context become plotter(newPs)

    {
       case Collaborate(group) =>
         update(copy(collaborators = collaborators ++ (group - self)))
       case Subordinates(recruits) =>
         recruits foreach { recruit => recruit ! Recruit }
         update(copy(subordinates = subordinates ++ recruits))
       case Recruit =>
         update(copy(superiors = superiors + sender))
       case UpdateProb(changes) =>
         update(copy(probs = probs ++ changes))
       case _ => // ignore other messages
    }
  }
}
\end{code}

Now we address how the plotters plot: if they receive a \texttt{Bad} message from within the covert group, they have some probability of continuing plot activity with their peers, some probability of delegating some work to subordinates (if they have any), and some probability of reporting to their supervisors (again if they have any).  We will use this as the probability of {\em any} \texttt{Bad} interaction with those groups, and that in general their interaction with these groups is binomially distributed.  Therefore, if a particular plotter has $k$ collaborators (or supervisors, or subordinates), the probability of interacting with an individual in that group, $p_i$, is related to the probability of interacting with the group, $p_g$ -- which is the probability in the simulation state -- by:

$$
(1-p_i)^k = 1 - p_g \rightarrow p_i = 1-\sqrt[k]{1-p_g}
$$

### 2014-01-28 rscripts/manuscript.tex
more updates

% PLOS One TEMPLATE
\usepackage{amsmath,amssymb,graphicx,cite,setspace}
\usepackage[textwidth=1.75in]{todonotes}
\usepackage[total={6in,9.5in},top=0.5in,left=0.5in,includefoot]{geometry}
\usepackage{listings}
\usepackage{url}
\usepackage{color}

% define Scala
\lstdefinelanguage{scala}{morekeywords={class,object,trait,extends,with,new,if,while,for,def,val,var,this,case,->,=>},
%otherkeywords={->,=>},
%procnamekeys={def,.},
sensitive=true,
keywordstyle=\color{magenta},
commentstyle=\color{green},
identifierstyle=\color{blue},
%procnamestyle=\color{black},
morecomment=[l]{//},
morecomment=[s]{/*}{*/},
morestring=[b]"}
% Default settings for code listings
\lstset{frame=tb,language=scala,aboveskip=3mm,belowskip=3mm,
showstringspaces=false,columns=flexible,basicstyle={\small\ttfamily}}
% handle tilde characters
\lstset{literate=%
{~}{\url{~}}1
}
% define no break code listing
\lstnewenvironment{code}[1][]%
{
   \noindent
   \minipage{\linewidth} 
   \vspace{0.5\baselineskip}
   \lstset{basicstyle=\ttfamily\footnotesize,frame=single,#1}}
{\endminipage}

@@

\usepackage{Sweave}
\begin{document}
\input{manuscript-concordance}
\begin{flushleft}
{\Large
\textbf{Modeling Requirements for Simulating Covert Social Groups}
}
% Insert Author names, affiliations and corresponding author email.
\\
Carl A. B. Pearson$^{1,\ast}$, 
Edo Airoldi$^{2}$,
Edward Kao$^{2}$,
Burton Singer$^{1}$, 
\\
\bf{1} Emerging Pathogens Institute, University of Florida, Gainesville, FL, USA
\\
\bf{2} Statistics, Harvard University, Cambridge, MA, USA
\\
$\ast$ E-mail: cap10@ufl.edu
\end{flushleft}
% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Covert social groups are difficult to observe directly, by definition: members either actively thwart discovery or leverage passive features of their surrounding population to avoid detection.  Moreover, increasing active observation would be expected to perturb these social systems.  However, these groups never have perfect cover, their secrecy may make them conspicuously absent from the everyday activity of the background population, and finally there are potential mechanical charactericterizations -- their composition, goals, tactics, etc. -- that might predict their actions.

This situation -- strong constraints on passively and actively obtained direct empirical data, overwhelming indirect data, and competing mechanical explanations -- recommends simulation.  However, because the of the censoring issues in the data, that simulation requires particular attention to characterizing uncertainty and avoiding overfitting.

Herein, we discuss these issues in terms of network models.  Specifically, we review (i) incorporating observation uncertainty, (ii) the need for both back- and foreground populations, (iii) avoiding over-fitting, and (iv) finally some philosophy relative to implementation.

As part of that philosophical discussion, we walkthrough the exercise of procedurally generating back- and foreground populations, simulating communication among those individuals, and filtering those communications via an observation model.  We then address the issues of (i) fitting that model against data, and (ii) analyzing performance of the opposing sides (e.g. Receiver Operator Characteristic).  Finally, we also demonstrate a small extension.

@@

One syntactical aside for the code the below sections: the Scala collections (like \texttt{Set}) define methods for adding and removing elements in terms of the natural mathematical operators (\texttt{+} and \texttt{-}) and for similar operations with groups of elements (\texttt{++} and \texttt{--}).  Similarly, the names for message senders (\texttt{sender}), changing context (\texttt{context become ...}), et cetera are all quite clear.  Given the informative, mixed-natural-and-mathematical-like syntax, we use actual code snippets instead of pseudo-code to illustrate.

\subsection*{The Covert Group}
First, we define the covert group context:

\begin{code}
  def plotter(
    superiors:Set[Plotter] = Set.empty,
    collaborators:Set[Plotter] = Set.empty,
    subordinates:Set[Plotter] = Set.empty) : Receive = {
      case _ => // do nothing with messages
  }
\end{code}

We need to define how they adopt and evolve this context.  Since we are not studying the organizational dynamics with this simulation, we can be phenomenologically loose for this part, though be taking this approach we have laid the ground for monitoring that activity as part of future detection scenarios.  The simulation runner will send a message \texttt{Radicalize} to the initial cabal, containing all its members:

\begin{code}
  case class Radicalize(id:Long, leaders:Set[Plotter])

  def receive = {
    case Radicalize(_, cabal) =>
      context become plotter(collaborators = (cabal - self))
    case _ => // do nothing with other messages
  }
\end{code}

@@

\begin{code}
  case class Subordinates(id:Long, recruits:Set[Plotter])
  case class Recruit(id:Long)
  case class RecruitAck(id:Long) // acknowledge recruitment

  def plotter(...) = {
    case Subordinates(_, recruits) =>
      recruits foreach { recruit =>
        recruit ! Recruit(0)
      }
    case RecruitAck(_) =>
      context become plotter(superiors, collaborators, subordinates + sender)
    // ... other cases
  }

  def receive = {
    case Recruit(id) =>
      context become plotter(superiors = Set(sender))
      sender ! RecruitAck(id)
    // ... other cases
  }
\end{code}

Some of these new recruits might be introduced to others as collaborators, so we need to need to handle collaboration messages, again assuming that the simulation runner is setting the particular covert structure.

\begin{code}
  case class Collaboration(id:Long, members:Set[Plotter])

  def plotter(...) = {
    case Collaboration(_, members) =>
      context become plotter(superiors, collaborators ++ members, subordinates)
    // ... other cases
  }
\end{code}

Similarly, we can advise these recruits about other actors for them to recruit.

Then we need to consider how these actors will know about each other in the simulation, how they will interact, etc.  In this example, we consider a static group, and as such the simulation could simply initialize the group.  However, we know that these groups do not really appear fully formed, and thus someday it will likely be desirable to include recruitment in the model.  We can anticipate that day by building our static initialization in those terms.

We will represent recruitment activity with two messages, which we add to the definition of the Leader:

\begin{code}[title=Covert Leader Helper Object]
object Leader {
  case class Recruit(id:Long) // try to recuit a member
  case class RecruitAck(id:Long) // join the group
}
\end{code}

We then modify the definitions of Leader and Member to respond to these messages:

\begin{code}[title=Using Messages]
class Leader extends Actor {
  var recruits = Set.empty[Member]

  def receive = {
    case RecruitAck(_) =>
      recruits += sender
    case _ =>
  }
}

class Member extends Actor {
  var leader;

  def receive = {
    case Recruit(id) =>
      leader = sender
    case _ =>
  }

}
\end{code}

>>>>

\usepackage{amsmath,amssymb,graphicx,cite,setspace} % PLOS One TEMPLATE
\usepackage{listings, url, color, appendix} % things I usually use
% items to make todonotes useful
\usepackage[textwidth=1.75in]{todonotes} 
\usepackage[total={6in,9.5in},top=0.5in,left=0.5in,includefoot]{geometry}
\newcommand{\wantFig}[2]{
  \begin{figure}[h]\centering
    \missingfigure{}
    \caption{#2}
    \figlabel{#1}
  \end{figure}
}
% alt layout formatting
%\doublespacing

% define Scala
\lstdefinelanguage{scala}{morekeywords={package,class,object,trait,extends,with,new,if,while,for,def,val,var,this,case,import,type,->,=>},
%otherkeywords={->,=>},
%procnamekeys={def,.},
sensitive=true,
keywordstyle=\color{magenta},
commentstyle=\color{green},
identifierstyle=\color{blue},
%procnamestyle=\color{black},
morecomment=[l]{//},
morecomment=[s]{/*}{*/},
morestring=[b]"}
% Default settings for code listings
\lstset{frame=tb,language=scala,aboveskip=3mm,belowskip=3mm,
showstringspaces=false,columns=flexible,basicstyle={\small\ttfamily}}
% handle tilde characters
\lstset{literate=%
{~}{\url{~}}1
}
% define no break code listing
\lstnewenvironment{code}[1][]%
{
   \noindent
   \minipage{\linewidth} 
   \vspace{0.5\baselineskip}
   \lstset{basicstyle=\ttfamily\footnotesize,frame=single,#1}}
{\endminipage}

@@

\newcommand{\figlabel}[1]{\label{fig_#1}}
\newcommand{\figref}[1]{fig. \ref{fig_#1}}
\newcommand{\Figref}[1]{Fig. \ref{fig_#1}}

\newcommand{\applabel}[1]{\label{app_#1}}
\newcommand{\appref}[1]{appendix \ref{app_#1}}
\newcommand{\Appref}[1]{Appendix \ref{fig_#1}}

\usepackage{Sweave}
\begin{document}
\input{manuscript-concordance}
\begin{flushleft}
{\Large
\textbf{Modeling Requirements for Simulating Covert Social Groups}
}
% Insert Author names, affiliations and corresponding author email.
\\
Carl A. B. Pearson$^{1,\ast}$, 
Edo Airoldi$^{2}$,
Edward Kao$^{2}$,
Burton Singer$^{1}$, 
\\
\bf{1} Emerging Pathogens Institute, University of Florida, Gainesville, FL, USA
\\
\bf{2} Statistics, Harvard University, Cambridge, MA, USA
\\
$\ast$ E-mail: cap10 \{at\} ufl \{dot\} edu
\end{flushleft}
% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Covert social groups are difficult to observe directly, by definition: members either actively thwart discovery or leverage passive features of their surrounding population to avoid detection.  Moreover, increasing active observation would be expected to perturb these social systems.  However, these groups never have perfect cover.  Their secrecy may make them conspicuously absent from the everyday activity of the background population.  Finally there are potential mechanical charactericterizations -- their composition, goals, tactics, etc. -- that might predict their actions.

The structure of covert networks is highly variable from one context to another, and they change composition and interaction patterns over time.  With any covert network, there are strong constraints on passively and actively obtainable direct empirical data, overwhelming amounts of indirect data, and competing mechanical explanations for what is observed.  Studying the detailed history of any one network, or set of them, will not necessarily provide insight about how to effectively detect such networks in the future.  This situation recommends simulation of a variety of scenarios with accompanying detection strategies. Because of the censoring issues in the data, simulation requires particular attention to characterizing uncertainty and avoiding overfitting.

Herein, we discuss these issues in terms of network models.  Specifically, we review (i) incorporating observation uncertainty, (ii) the need for both back- and foreground populations, (iii) avoiding over-fitting, and (iv) some philosophy relative to implementation.

As part of that philosophical discussion, we walkthrough the exercise of procedurally generating back- and foreground populations, simulating communication among those individuals, and filtering those communications via an observation model.  We then address the issues of (i) fitting that model against data, and (ii) analyzing performance of the opposing sides (e.g. Receiver Operator Characteristic).  Finally, we also demonstrate a small model extensionand show how it affects the (unextended) detection performance.

@@

One syntactical aside for the code the below sections: the Scala collections (like \texttt{Set}) define methods for adding and removing elements in terms of the natural mathematical operators (\texttt{+}, \texttt{-}) and for similar operations with groups of elements (\texttt{++}, \texttt{--}).  Similarly, the names for message senders (\texttt{sender}), changing context (\texttt{context become ...}), et cetera are all quite clear.  Given the informative, mixed-natural-and-mathematical-like syntax, we prefer actual code snippets to pseudo-code to illustrate simulation steps.  However, we have elided some declarations that enable the more informative syntax from the initial appearance in code.

\subsection*{The Covert Group}
First, we define the covert group \texttt{Receive} context:

\begin{code}
  def plotter(
    // a plotter has superiors, collaborators, and subordinates
    // these groups are initially empty
    superiors:Plotters = empty,
    collaborators:Plotters = empty,
    subordinates:Plotters = empty
    /*, ... other args */) : Receive = {
      // ... to-be-defined behavior
      case _ => // ignore unmatched messages
  }
\end{code}

We need to define how they adopt and evolve this context.  Since we are not studying the organizational dynamics with this simulation, we can be phenomenologically loose for this part, though this approach means we can add monitoring that activity as part of future detection scenarios with minimal coding.  The simulation runner will send a \texttt{Radicalize} message to the initial cabal, then a \texttt{Collaborate} message to the same group:

\begin{code}
  object Radicalize
  case class Collaborate(group:Plotters)

  def receive = {
    case Radicalize =>
      context become plotter
    case _ => // ignore other messages
  }

  def plotter(...) => {
    case Collaboration(_, group) =>
      context become plotter(collaborators = collaborators ++ (group - self))
    // ...
  }
\end{code}

@@

\begin{code}
  case class Subordinates(recruits:Plotters)
  object Recruit

  def plotter(...) = {
    case Subordinates(_, recruits) =>
      recruits foreach { recruit =>
        recruit ! Recruit
      }
    // ... other cases
  }

  def receive = {
    case Recruit =>
      context become plotter(superiors = Set(sender))
    // ... other cases
  }
\end{code}

At this point, we have all the necessary messages to construct the given network.  Those messages, based on the labels in Sageman et al. are covered in \Appref{sage_network}.  However, we still need to model how the plotters actually plot.  We will assume that all plotting (i) originates with leadership, (ii) when intermediates receive direction, they work with their \texttt{collaborators} as well as delegating work, (iii) that directions occasionally result in feedback to higher-ups, and finally (iv) all of the ``work'' and direction is captured as passing \texttt{Bad} messages.

The simulation runner sends \texttt{Time(t:Long)} messages to all actors in the simulation to indicate the march of time (waiting for them to complete all their business for that step and reply with \texttt{TimeAck(t)}).  In general, during each time step a leader may initiate a new \texttt{Bad} message.  Other members of the covert group circulate \texttt{Bad} messages subsequent to receiving a \texttt{Bad} message themselves, with different probabilities for sending messages to superiors, collaborators, and subordinates.  Those probabilities do not depend on the origin of the received message (though this could be a useful mechanical tweak to the model).

Though those probabilities could be coded into the actors as part of their initialization, if they are set and adjusted by events then the actors (i) more accurrately reflect real behavior and (ii) the evolution of individuals occurs via a consistent simulation mechanism.  The former provides a lever for expert insight and empirical study.  The latter makes it easier to produce real-alike data -- i.e., simulation data that has the form of real observations.

In our simplified example, we do not consider how ``real'' events would translate into adjusting a covert member's behavior.  Instead, the simulation runner sends a message, giving us a wedge for later adjustment without frontloading too much model complexity.  This represents the final step for how the plotters evolve in this simulation, so the following repeats the previous snippets with some condensing modifications to the code.  The most notable one is making use of a \texttt{case class} to represent all of the parameters that form a plotter's state.

\begin{code}
package object simactor { // defs to make implementation more natural
  type Plotters = Set[ActorRef]
  def empty : Plotters = Set.empty
  def group(list:ActorRef*) : Plotters = list.toSet
  type Probability = Double
}

object SimActor { // defines common elements for SimActors
  // MESSAGES ABOUT INTERNAL STATE
  object Radicalize                           // become a plotter
  object Recruit                              // work for a plotter
  case class Collaborate(group:Plotters)      // work with other plotters
  case class Subordinates(recruits:Plotters)  // recruit subordinates

  object PType extends Enumeration { // plotter behavior probabilites
    type PType = Value
    val Initiate, // prob. of starting plot comms
        Report,   // prob. of reporting to superiors
    	  Collab,	  // prob. of working w/ peers
    	  Delegate  // prob. of delegating work
    		  = Value
  }
  case class UpdateProb(changes:Map[PType.Value,Probability])

  // MESSAGES ABOUT EXTERNAL ACTIVITY
  object Bad
  object Good

  case class PlotterState( // the plotter state description
    superiors:Plotters = empty,
    collaborators:Plotters = empty,
    subordinates:Plotters = empty,
    probs:Map[PType.Value,Probability] = PType.values.map { (_, 0.0) } toMap
    // default each plotting probability 0.0
  )
}

class SimActor extends Actor {
  import SimActor._

  def receive = {
    case Radicalize =>
      context become plotter()
    case Recruit =>
      context become plotter(PlotterState(superiors = group(sender)))
    case _ => // ignore other messages
  }

  def plotter(ps:PlotterState = PlotterState()) : Receive = {
    import ps._
    def update(newPs:PlotterState) = context become plotter(newPs)

    {
       case Collaborate(group) =>
         update(copy(collaborators = collaborators ++ (group - self)))
       case Subordinates(recruits) =>
         recruits foreach { recruit => recruit ! Recruit }
         update(copy(subordinates = subordinates ++ recruits))
       case Recruit =>
         update(copy(superiors = superiors + sender))
       case UpdateProb(changes) =>
         update(copy(probs = probs ++ changes))
       case _ => // ignore other messages
    }
  }
}
\end{code}

Now we address how the plotters plot: if they receive a \texttt{Bad} message from within the covert group, they have some probability of continuing plot activity with their peers, some probability of delegating some work to subordinates (if they have any), and some probability of reporting to their supervisors (again if they have any).  We will use this as the probability of {\em any} \texttt{Bad} interaction with those groups, and that in general their interaction with these groups is binomially distributed.  Therefore, if a particular plotter has $k$ collaborators (or supervisors, or subordinates), the probability of interacting with an individual in that group, $p_i$, is related to the probability of interacting with the group, $p_g$ -- which is the probability in the simulation state -- by:

$$
(1-p_i)^k = 1 - p_g \rightarrow p_i = 1-\sqrt[k]{1-p_g}
$$
